---
title: "Disambiguation Database"
author: "Akila Forde"
date: '2022-07-21'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(data.table)

#setwd( "H:/data-raw/bmf"  )
```


<!-- ## Testing Disambiguation Database -->

<!-- First, getting a test file from the NCCS Business Master File [resource](https://datacatalog.urban.org/dataset/national-center-charitable-statistics-ncc-irs-business-master-file/resource/ffbeb4d8-27e5).  -->

<!-- Figure out if this is only for a single year -->

<!-- ```{r} -->
<!-- # download all regions and combine into a single dataset -->

<!-- region.01 <- read.csv( "https://www.irs.gov/pub/irs-soi/eo1.csv" , stringsAsFactors=F )   -->

<!-- region.02 <- read.csv( "https://www.irs.gov/pub/irs-soi/eo2.csv" , stringsAsFactors=F ) -->

<!-- region.03 <- read.csv( "https://www.irs.gov/pub/irs-soi/eo3.csv" , stringsAsFactors=F ) -->

<!-- region.04 <- read.csv( "https://www.irs.gov/pub/irs-soi/eo4.csv" , stringsAsFactors=F ) -->


<!-- bmf <- rbind( region.01, region.02, region.03, region.04 ) -->

<!-- rm(region.01, region.02, region.03, region.04) -->

<!-- ``` -->


<!-- ```{r} -->

<!-- # The rule date is formatted as YYYY-MM, so split out the year -->

<!-- bmf$RULEYEAR <- substr( bmf$RULING, 1, 4 ) -->

<!-- # barplot of ruling dates -->
<!-- options( scipen=8 ) -->
<!-- barplot( table( bmf$RULEYEAR ), las=2, cex.names=0.5, col="gray40", border=NA ) -->
<!-- abline( h=seq(20000,100000,20000), col="white", lwd=2 ) -->
<!-- title( main="Ruling Dates of Existing Nonprofits" ) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- # the NTEE codes created by the IRS are very messy with lots of errors -->

<!-- #table( bmf$NTEE ) -->

<!-- # create to top level group - see documentation below -->

<!-- bmf$LEVEL4 <- substr( bmf$NTEE_CD, 0, 1 ) -->


<!-- ``` -->


<!-- ```{r} -->
<!-- # SEPARATE ACTIVITY CODES -->
<!-- # -->
<!-- # The variable ACTIVITIES consists of three separate 3-digit activity codes. -->
<!-- # Split these into separate variables. -->
<!-- # Documentation available below. -->

<!-- bmf$ACT1 <- substr( bmf$ACTIVITY, 1, 3 ) -->

<!-- bmf$ACT2 <- substr( bmf$ACTIVITY, 4, 6 ) -->

<!-- bmf$ACT3 <- substr( bmf$ACTIVITY, 7, 9 ) -->
<!-- ``` -->



<!-- # Cleaning the bmf data -->

<!-- To prepare the Business Master File, create a file with (1) name, (2) ein, (3) address fields, (4) NTEE code, (5) size (expenses, revenue, assets), (6) source and date of data. -->

<!-- ```{r} -->

<!-- bmf_file <- bmf %>% mutate(ZIP5 = substring(ZIP, 1, 5)) %>% -->
<!--             select(NAME, EIN, STREET, CITY, STATE, ZIP5, NTEE_CD, -->
<!--                    ASSET_AMT, INCOME_AMT, REVENUE_AMT, SORT_NAME) -->


<!-- saveRDS( bmf_file, file="BusinessMasterFile.rds" ) -->
<!-- ``` -->


<!-- # Testing matching -->

<!-- ```{r} -->
<!-- library(reclin2) -->

<!-- #Load saved BMF file -->
<!-- bmf <- readRDS("BusinessMasterFile.rds") -->

<!-- set.seed(17) -->

<!-- #Load organization input -->
<!-- yes.match <- read.csv("../data-raw/Matches1to13615.csv") %>% -->
<!--         select( recipient_name, recipient_city_name,  -->
<!--               recipient_state_code, recipient_zip) %>% -->
<!--   rename( "NAME" = "recipient_name" , "CITY" = "recipient_city_name" , -->
<!--          "STATE" = "recipient_state_code", "ZIP5" = "recipient_zip" ) %>%  -->
<!--   #Convert to uppercase -->
<!--   mutate_if(is.character, str_to_upper) %>%  -->
<!--   #Pad Zipcode with zeros if string is less than 5 characters -->
<!--   mutate(ZIP5 = str_pad(ZIP5, 5, side="left", pad="0")) -->

<!-- #Taking a sample -->
<!-- yes.sample <- sample_n(yes.match, 100) -->



<!-- ``` -->

<!-- Blocking on State, returns 6.3 million pairs (100 inputs with 1.8 mil in the BMF) and matching on Largest common substring for both Name and City. -->

<!-- ```{r} -->
<!-- y.pairs <-  pair_blocking(yes.sample, bmf, "STATE") %>%  -->
<!--           compare_pairs(y.pairs, on = c("NAME", "CITY"),  -->
<!--                         comparators = list(NAME = lcs(), CITY = lcs())) -->


<!-- model <- problink_em(~ NAME, data = y.pairs) -->

<!-- pairs <- predict(model, y.pairs, c("weights", "mpost"), add = TRUE, binary = TRUE) -->


<!-- pairs <- select_greedy(pairs, "weights", variable = "greedy", threshold = 0) -->


<!-- pairs <- select_n_to_m(pairs, "weights", variable = "ntom", threshold = 0) -->


<!-- ``` -->

<!-- Linking pairs based on name -->

<!-- ```{r} -->
<!-- greedy <- link(pairs, selection = "greedy") -->

<!-- # Impressive -->
<!-- ntom <- link(pairs, selection = "ntom") -->
<!-- ``` -->

<!-- NtoM matched 80/100 NPOs accurately, Greedy also matched 80/100 but had false positives (When matched on just name) -->


<!-- ```{r} -->

<!-- unmatched <- yes.sample$NAME[which(!(yes.sample$NAME %in% ntom$NAME.x))] -->
<!-- ``` -->


<!-- ```{r} -->
<!-- unmatched -->


<!-- ``` -->


# Testing BMF Master file

```{r}

#Load saved BMF file
bmf <- readRDS("../data-raw/bmf/bmf-master.rds")

bmf <- bmf %>% 
  mutate(ZIP5 = str_pad(ZIP5, 5, side="left", pad="0")) %>% 
            select(NAME, EIN, ADDRESS, CITY, STATE, ZIP5, NTEEFINAL, 
                   ASSETS, INCOME)


```

Load in the test data, converting case to upper and padding the zipcode with extra zeros if too short. Then taking a sample of 100 organizations. For purposes of user use, the organizations should be in the form of a dataframe if components has more than just org name.

```{r}
set.seed(17)

#Load organization input
yes.match <- read.csv("../data-raw/Matches1to13615.csv") %>%
        select( recipient_name, recipient_city_name, 
              recipient_state_code, recipient_zip, EIN) %>%
  rename( "NAME" = "recipient_name" , "CITY" = "recipient_city_name" ,
         "STATE" = "recipient_state_code", "ZIP5" = "recipient_zip" ) %>% 
  #Convert to uppercase
  mutate_if(is.character, str_to_upper) %>% 
  #Pad Zipcode with zeros if string is less than 5 characters
  mutate(ZIP5 = str_pad(ZIP5, 5, side="left", pad="0"))

#Taking a sample
yes.sample <- sample_n(yes.match, 100)

rm(yes.match)

```

## If input organization has address attributes like State, City:

Blocking on state to speed it up

```{r}
library(reclin2)

system.time(y.pairs <-  pair_blocking(yes.sample, bmf, "STATE") %>% 
          compare_pairs(y.pairs, on = c("NAME", "CITY"), 
                        comparators = list(NAME = lcs(), CITY = lcs())))


system.time(model <- problink_em(~ NAME, data = y.pairs))

system.time(pairs <- predict(model, y.pairs, c("weights", "mpost"), add = TRUE, binary = TRUE))


system.time(pairs <- select_greedy(pairs, "weights", variable = "greedy", threshold = 0))


system.time(pairs <- select_n_to_m(pairs, "weights", variable = "ntom", threshold = 0))


```
Returns 89 matches, but with more false positives on the ntom side


```{r}
greedy <- link(pairs, selection = "greedy")

# Impressive
ntom <- link(pairs, selection = "ntom")
```


# Find the non-matches


```{r}
# CHECKING THOSE THAT ARE IN THE BMF

length( intersect( yes.sample$EIN, bmf$EIN ) )

setdiff(ntom$EIN.y, yes.sample$EIN)

setdiff(ntom$EIN.y, ntom$EIN.x)


### Returning the pairs that didn't match
false_negs <- pairs$.x[which(pairs$.x == pairs$.y & pairs$ntom == FALSE)]
false_negs_y <- pairs$.y[which(pairs$.x == pairs$.y & pairs$ntom == FALSE)]


pairs[which(pairs$.x == pairs$.y & pairs$ntom == FALSE)]
yes.sample[false_negs,]
bmf[false_negs_y,]


false_pos <- pairs$.x[which(pairs$.x != pairs$.y & pairs$ntom == TRUE)]
false_pos_y <- pairs$.y[which(pairs$.x != pairs$.y & pairs$greedy == TRUE)]


```



#Parallelization of reclin2

Excellent tutorial found here: https://cran.r-project.org/web/packages/reclin2/vignettes/using_a_cluster_for_record_linkage.html

```{r}
library(parallel)
#library(reclin2)

num.cores <- detectCores()

#Start a cluster with n nodes.
cl <- makeCluster(num.cores - 1 )

#Execute any pre-processing code necessary in each node (e.g. loading a package)cl <- makeCluster(2)


#Blocking pairs by State, if address is included
cl.pairs <- cluster_pair_blocking(cl, yes.sample, bmf, "STATE")
#system.time(cl.pairs <- cluster_pair_blocking(cl, yes.sample, bmf, "STATE"))


# Collect pairs from the clusters
#local_pairs <- cluster_collect(cl.pairs, clear = FALSE)
system.time(local_pairs <- cluster_collect(cl.pairs, clear = FALSE))

## Compare names by Longest common substring algorithm
#compare_pairs(cl.pairs, on = "NAME", comparators = list(NAME = lcs()))
system.time(compare_pairs(cl.pairs, on = "NAME", comparators = list(NAME = lcs())))

## Implement model based on Name 
#cl.model <- problink_em(~ NAME, data = cl.pairs)
system.time(cl.model <- problink_em(~ NAME, data = cl.pairs))


## Predict weights based on Name
#cl.pairs <- predict(cl.model, cl.pairs, type = "weights", add = TRUE, binary = TRUE)
system.time(cl.pairs <- predict(cl.model, cl.pairs, type = "weights", add = TRUE, binary = TRUE))

# Select and collect pairs which meet a certain threshold
system.time(cl.pairs <- select_threshold(cl.pairs, "threshold", score = "weights", threshold = 0))
system.time(local_pairs <- cluster_collect(cl.pairs, "threshold"))


# Select pairs using the ntom algorithm
system.time(local_pairs <- select_n_to_m(local_pairs, "weights", variable = "ntom", threshold = 0))


# True matched pairs
local_pairs <- compare_vars(local_pairs, "truth",  on_x = "EIN", on_y = "EIN")

table(local_pairs$truth, local_pairs$ntom)

system.time(linked_data_set <- link(local_pairs, selection = "ntom"))

stopCluster(cl)


### Will need to re-order the items
linked_data_set <- cbind(yes.sample[linked_data_set$.x, ] , bmf[linked_data_set$.y,] )

```


## To-do

Run agrep version - to gain time for

```{r}
system.time(
for (i in nrow(yes.sample)){
  for (j in nrow(bmf)){
    
    if (agrep(yes.sample$NAME[i], bmf$NAME[j], max.distance = 0.2) != 0){
      print(yes.sample$NAME[i], bmf$NAME[j] )
    }
    
  }
}
)

```



# Change the clusters


<!-- Introducing the fuzzy wuzzy package to better refine the matches from reclin2 -->

<!-- ```{r} -->
<!-- # Must load reticulate::py_config() before loading the package -->
<!-- reticulate::py_config() -->

<!-- library(fuzzywuzzyR) -->

<!-- refine_matches <- function(df){ -->

<!--   dt <- data.frame() -->
<!--   for (i in 1:nrow(df)) { -->

<!--     # returns ratio of two strings potentially matching, uses python's diflib package -->
<!--     init = SequenceMatcher$new(string1 = df$NAME.x[i], string2 = df$NAME.y[i]) -->

<!--     if (init$ratio() > 0.8){ -->

<!--       dt <- rbind(dt, df[i,]) -->
<!--     } -->

<!--   } -->
<!--   return(dt) -->

<!-- } -->

<!-- ``` -->



<!-- # Testing fuzzyjoin package -->
<!-- # ```{r} -->
<!-- #  -->
<!-- # library(fuzzyjoin) -->
<!-- #  -->
<!-- # fuzz.join <- stringdist_inner_join(yes.sample, bmf, -->
<!-- #                       method = "lv", -->
<!-- #                       by = c("NAME", "STATE"), -->
<!-- #                       max_dist = 6) -->
<!-- # ``` -->

